\vspace{-2cm} \LARGE Divergence-Free Smoothed Particle Hydrodynamics \\ \large Advanced Simulation and Visualization of Fluids in Computer Graphics

### 2016-12-14 report/paper.tex
density solver

    \begin{equation} \label{eq:nsVelocity}
        \frac{D\textbf{v}}{Dt} = \frac{1}{\rho} \bigtriangledown p + v \bigtriangledown ^2 \textbf{v} + \frac{\textbf{f}}{\rho}
    \end{equation}

    In the equations $\frac{D\rho}{Dt}$ and $\frac{D\textbf{v}}{Dt}$ denotes the derivate of the density $\rho$ and the velocity $\textbf{v}$ respectively. The pressure, kinematic viscosity and body forces are denoted by $p,v$ and $\textbf{f}$ respectively. The second part of Equation~\ref{eq:nsDensity} $\bigtriangledown \cdot \textbf{v} = 0$ induce that the velocity field is free from divergence. From the continuity equation $\frac{D\rho}{Dt} = -\rho \bigtriangledown \cdot \textbf{v}$ and the divergence-free condition it follows that the partial derivate of the density with respect to the time is zero, which implies the equivalence of the equations in Equation~\ref{eq:nsDensity}. From this is follows that the density, in theory, must stay constant over time which is the implication of an incompressible fluid. In practice the divergence-free condition is not sufficient to guarantee incompressibility in simulations. The numerical time integrations will contain numerical errors which will cause density deviations due to the volume compressions. To avoid this a second condition $\rho - \rho_0 = 0$ called \textit{constant density condition} must be fullfilled~\cite{bender}. Section~\ref{divergenceSolver} and~\ref{densitySolver} give a deeper description of the solvers used in the project.

\subsection{Divergence solver} \label{divergenceSolver}
\subsection{Density solver} \label{densitySolver}

>>>>

    \begin{equation} \label{eq:nsVelocity}
        \frac{D\textbf{v}}{Dt} = -\frac{1}{\rho} \bigtriangledown p + v \bigtriangledown ^2 \textbf{v} + \frac{\textbf{f}}{\rho}
    \end{equation}

    In the equations $\frac{D\rho}{Dt}$ and $\frac{D\textbf{v}}{Dt}$ denotes the derivate of the density $\rho$ and the velocity $\textbf{v}$ respectively. The pressure, kinematic viscosity and body forces are denoted by $p,v$ and $\textbf{f}$ respectively. The second part of Equation~\ref{eq:nsDensity} $\bigtriangledown \cdot \textbf{v} = 0$ induce that the velocity field is free from divergence. From the continuity equation $\frac{D\rho}{Dt} = -\rho \bigtriangledown \cdot \textbf{v}$ and the divergence-free condition it follows that the partial derivate of the density with respect to the time is zero, which implies the equivalence of the equations in Equation~\ref{eq:nsDensity}. From this is follows that the density, in theory, must stay constant over time which is the implication of an incompressible fluid. In practice the divergence-free condition is not sufficient to guarantee incompressibility in simulations. The numerical time integrations will contain numerical errors which will cause density deviations due to the volume compressions. To avoid this a second condition $\rho - \rho_0 = 0$ called \textit{constant density condition} must be fullfilled~\cite{bender}. Bridson~\cite{bridson} uses splitting on Equation~\ref{eq:nsVelocity} which induce $\frac{D\textbf{v}}{Dt} = -\frac{1}{\rho} \bigtriangledown p$ which will be used further on in the simulation. Section~\ref{divergenceSolver} and~\ref{densitySolver} give a deeper description of the solvers used in the project.

\subsection{Divergence solver} \label{divergenceSolver}
\subsection{Density solver} \label{densitySolver}
    This solver aims to minimize the density error caused by the numerical integration in the divergence solver. The density error is determined by comparing the actual density to the rest density, $\rho - \rho_0$. There are several density solvers presented in previous works, but the solver used in this project was presented by Bender et. al.~\cite{bender} which uses the fact that $\alpha_i$ has been calculated before the solver which reduces the computations in the solver. The solver will be active while the average of the predicted densities for every particle differs from the rest density with an error greater than a constant \textit{maxError}. For every particle a predicted density $\rho_i^*$ is calculated through an Euler integration step by Equation~\ref{eq:predDensity}.

    \begin{equation} \label{eq:predDensity}
        \resizebox{0.88\linewidth}{!}{$
        \begin{matrix}
        \rho_i^* = \rho_i + \Delta t \frac{D\rho_i}{Dt} \\ 
        \\
        \Leftrightarrow \\ 
        \\
        \rho_i^* = \rho_i + \Delta t + \sum_j m_j 
        (\textbf{v}_i^* - \textbf{v}_j^*) \bigtriangledown W_{ij}
        \end{matrix}
        $}
    \end{equation}

    Since the stiffness parameter $\kappa_i$ is derived from the pressure forces, updating the stiffness parameter will correct the density error. The stiffness parameter is determined by Equation~\ref{stiffnessDensity}.

    \begin{equation} \label{stiffnessDensity}
    \kappa_i = \frac{1}{\Delta t^2}(\rho_i^*-\rho_0)\alpha_i
    \end{equation}

    The velocity is updated according to the the divergence solver. %Det här behöver kanske utvecklas, väntar med det tills stycket om divergence solvern är klar, innehåller samma ekvationer delvis
